---
title: "Stastical Case Studies Project 1 part 1"
author: "Georgia Ettles, Mayez Haris and Niharika Peddinenikalva"
date: '2022-10-28'
output:
  pdf_document: default
  html_document: default
---

This project had equal contributions from all members. All 3 members sorted and cleaned the data together along with building the model. Mayez and Georgia refined the models to choose the most suitable one. Niharika wrote the code for box plots and other plots in the project. Georgia and Mayez wrote the analysis of the models while Niharika defined the model and formatted the document.

\newpage
# Introduction

The easySHARE data (https://tinyurl.com/wud2ynbs) is in the form of panel data. This report aims to analyse the data and investigate which risk factors have a significant impact on the participants’ BMI, and how these risk factors may affect men and women differently. The data was collected in different waves at different points in time, leading to some individuals participating in the study multiple times, and the questions asked were slightly different for each wave. To reduce repeating entries and to simplify the data that was analysed, this report focuses on wave 5 carried out in four countries - Austria, Germany, The Netherlands, and Sweden - chosen arbitrarily.

The data has 18749 observations, with 8479 men and 10270 women, and the average participants are aged 66 (since the youngest participant was aged 30). The interviews for cities and suburbs had 28.15% of participants.  32.45% participants have graduated secondary school, 80.5% of the participants are able to make ends meet and 94.45% of the participants live in a household of 4 or fewer members. Also, 66.56% of the participants have 0-2 children. 97.37% participants have higher mobility according to the Activities of Daily living index while 57.99% frequently perform vigorous activities. The average BMI of the participants is 26.62. This is in the overweight range.


```{r setup, echo=FALSE, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
suppressPackageStartupMessages(library("tidyverse"))
suppressPackageStartupMessages(library("dplyr"))
suppressPackageStartupMessages(library("gridExtra"))
suppressPackageStartupMessages(library("visdat"))
suppressPackageStartupMessages(library("countrycode"))
suppressPackageStartupMessages(library("ggplot2"))
suppressPackageStartupMessages(library("knitr"))
```

```{r Load-data}
# Load and store data
load("easySHARE_rel8_0_0(1).rda")
data <- easySHARE_rel8_0_0
```

```{r fix-missing-data}
# Set missing values to NA       
data[data < 0] <- NA

# Replace country codes with country names
data$country_mod <- countrycode(factor(data$country_mod),
  origin = "iso3n",
  destination = "country.name"
)
```


# Defining Model and Checking Model Assumptions

We consider the model below to explore risk factors of obesity. Let, for $n$ observations, $x_{i,j}$ represent the $i = 1, 2, \dots, n$ observation of the variable $x_j$ in the table below. The risk factors are represented by the explanatory variables $x_j$ and described in Table 1 using the variable name on the easySHARE data guide.

\begin{align*}
E(Y) = & \beta_0 + \sum_{j=1}^{22} \beta_j x_j + \beta_{10,3} x_{10}x_3 + \beta_{5,1} x_5 x_1 + \beta_{2,15} x_2 x_{15} + \beta_{2,16} x_2 x_{16} \\
& + \beta_{2,19} x_2 x_{19} + \beta_{7,18} x_7 x_{18} + \beta_{7,20} x_7 x_{20} + \beta_{7,21} x_7 x_{21} \\
& + \beta_{12,16} x_{12} x_{16} + \beta_{14,19} x_{14} x_{19} + \beta_{16,1} x_{16} x_1 + \beta_{19,1} x_{19} x_1
\end{align*}

\[
\implies E(\mathbf{Y}) = X \boldsymbol{\beta}, \text{where}
\]
\[
X = \begin{pmatrix} 
1 & x_{1,1} & x_{1,2} & \dots & x_{1,22} & x_{1,10} x_{1,3} & x_{1,5} x_{1,1} & \dots & x_{1,16} x_{1,1} & x_{1,19} x_{1,1} \\
\vdots & \vdots & \vdots & \ddots & \vdots & \vdots & \vdots & \ddots & \vdots & \vdots \\
1 & x_{n,1} & x_{n,2} & \dots & x_{n,22} & x_{n,10} x_{n,3} & x_{n,5} x_{n,1} & \dots & x_{n,16} x_{n,1} & x_{n,19} x_{n,1}
\end{pmatrix} \text{ and } \quad 
\boldsymbol{\beta} = \begin{pmatrix} \beta_0 \\ \beta_1 \\ \vdots \\ \beta_{22} \\ \beta_{10,3} \\ \beta_{5,1} \\ \vdots \\ \beta_{16,1} \\ \beta_{19,1} \end{pmatrix}
\]



```{=latex}
\begin{table}[!h]
\centering
\begin{tabular}{|c|c|p{4cm}|p{6cm}|}
\hline
$x_j$ & \textbf{Variable name} & \textbf{Description of variable}  & \textbf{Possible Values}\\
\hline
$x_1$ & age & Age of participant & range of positive values \\
\hline
$x_2$ & female & Male or female &  Male=0, Female =1 \\
\hline
$x_3$ & dn004\_mod & Born in country of interview &  No=0, Yes=1 \\
\hline
$x_4$ & partnerinhh & Living with spouse/partner &  No=0, Yes=1 \\
\hline
$x_5$ & mother\_alive & Is mother alive? &  No=0, Yes=1 \\
\hline
$x_6$ & father\_alive & Is father alive? &  No=0, Yes=1 \\
\hline
$x_7$ & ever\_smoked & Ever smoked daily? &  No=0, Yes=1 \\
\hline
$x_8$ & hc012\_ & Visited hospital in last 12 months? &  No=0, Yes=1 \\
\hline
$x_9$ & hc029\_ & In a nursing home? &  No=0, Yes=1 \\
\hline
$x_{10}$ & country\_mod & Country identifier &  Name of each country* \\
\hline
$x_{11}$ & iv009\_mod & Area of interview location &  City\_sub = City or suburbs; Town\_rural = Small/Large towns or village \\
\hline
$x_{12}$ & hhsize & Household size &  Up to 4; More than 4 people \\
\hline
$x_{13}$ & eurod & Depression scale of EURO-D &  Not depressed = 0-3, Somewhat depressed = 4-6, Relatively depressed = 7-9, Very depressed = 10-12 \\
\hline
$x_{14}$ & ch001\_ & Number of children alive &  Up to 2 or More than 2 children  \\
\hline
$x_{15}$ & chronic\_mod & Number of chronic diseases &  0; 1-3; 4-10 chronic diseases   \\
\hline
$x_{16}$ & adla & Activities of daily living index &  High mobility = 0-2; Low mobility = 3-5\\
\hline
$x_{17}$ & br015\_ & Vigorous activities &  Often = Once a week or more; Rarely = Less than once a week \\
\hline
$x_{18}$ & co007\_ & Is household able to make ends meet? &  With difficulty = some or great difficulty; Relatively easily = fairly easily or easily \\
\hline
$x_{19}$ & br010\_mod & Drinking behaviour &  Rarely = Less than once a month or never; Occasionally = once a month to once or twice a week; Regularly = 3 days a week to almost every day \\
\hline
$x_{20}$ & maxgrip & Maximum of grip strength measure &  Low = Less than or equal to 35; High = greater than 35 \\
\hline
$x_{21}$ & sphus & Self-perceived health &  Good = Excellent/Very good/Good; Not good = Fair/Poor \\
\hline
$x_{22}$ & isced1997\_r & Level of Education completed (ISCED-97 coding) &  Secondary and below; Post secondary; other = Still in school/other \\
\hline

\end{tabular}
\caption{List of explanatory variables}
\end{table}
```
\newpage

Note: $x_2$ to $x_{22}$ are categorical variables while $x_1=$ age is a continuous variable.


```{r Set-lists}
# binary variables
factor_list <- c(
  "female", "dn004_mod",
  "partnerinhh", "mother_alive", "father_alive", "ever_smoked",
  "hc012_", "hc029_"
)

# variables with more than two levels
ordered_list <- c(
  "iv009_mod", "hhsize", "eurod",
  "ch001_", "chronic_mod", "adla", "br015_", "co007_",
  "br010_mod", "maxgrip", "sphus", "isced1997_r"
)

# continuous variables
nonfactor_list <- c("age", "country_mod")
```

```{r factor-levels}
# Factoring binary variables as dummy variables
for (column in factor_list) {
  if (column != "hc029_") {
    data[, column] <- ifelse(data[, column] == "1", 1, 0)
  } else {
    data$hc029_ <- ifelse(data[, column] == "5", 0, 1)
  }
  data[, column] <- factor(data[, column], levels = c(0, 1), labels = c("No", "Yes"))
}


# Factor non-binary variables
for (column in ordered_list) {
  data[, column] <- factor(data[, column])
}

# Simplifying levels for specific variables
levels(data$isced1997_r) <- list(
  "Secondary and Below" = as.character(0:3),
  "Post Secondary" = as.character(4:6),
  "Other" = as.character(95:96)
)

levels(data$eurod) <- list(
  "Not Depressed" = as.character(0:3),
  "Somewhat Depressed" = as.character(4:6),
  "Relatively Depressed" = as.character(7:9),
  "Very Depressed" = as.character(10:12)
)

levels(data$hhsize) <- list(
  "Up to 4" = as.character(1:4),
  "More than 4" = as.character(c(4:12, 14:15))
)


levels(data$iv009_mod) <- list(
  "City_sub" = as.character(1:2),
  "Town_rural" = as.character(3:5)
)

levels(data$sphus) <- list(
  "Good" = as.character(1:3),
  "Not Good" = as.character(4:5)
)

levels(data$ch001_) <- list(
  "Up to 2" = as.character(0:2),
  "More than 2" = as.character(3:19)
)

levels(data$chronic_mod) <- list(
  "0" = as.character(0),
  "1-3" = as.character(1:3),
  "4-10" = as.character(4:10)
)

levels(data$adla) <- list(
  "High Mobility" = as.character(0:2),
  "Low Mobility" = as.character(3:5)
)

levels(data$br015_) <- list(
  "Often" = as.character(1:2),
  "Rarely" = as.character(3:4)
)

levels(data$co007_) <- list(
  "With Difficulty" = as.character(1:2),
  "Relatively easily" = as.character(3:4)
)

levels(data$br010_mod) <- list(
  "Rarely" = as.character(1:2),
  "Occasionally" = as.character(3:4),
  "Regularly" = as.character(5:7)
)

levels(data$maxgrip) <- list(
  "Low" = as.character(1:35),
  "High" = as.character(36:99)
)
```


\newpage
Below is the summary of the data that was used for the analysis of the model.


```{r setting-up-data}
# Setting up data to analyse choosing the 5th wave and 4 countries
working_data <- data %>%
  filter(wave %in% c(5)) %>%
  filter(country_mod %in% c("Austria", "Germany", "Netherlands", "Sweden")) %>%
  select(all_of(c(factor_list, ordered_list, nonfactor_list)), "bmi")

# Setting factor levels for the countries in the data
working_data$country_mod <- factor(working_data$country_mod)

options(knitr.kable.NA = "")

# Display summary statistics of chosen data using knitr
knitr::kable(summary(working_data %>% select(all_of(factor_list))))
knitr::kable(summary(working_data %>% select(c("hhsize",  "sphus", 
                                               "ch001_",  "adla", "br015_"))))
knitr::kable(summary(working_data %>% select(c("br010_mod", "chronic_mod", "isced1997_r", "country_mod", "eurod"))))
knitr::kable(summary(working_data %>% select(c("co007_", "maxgrip", "iv009_mod",
                                               "age", "bmi")))) 
```


The data above contains `NA`s which were filtered out while constructing the male and female models. 

\newpage
```{r Box-Plots-countries, fig.height=7.5, fig.width=6.5, fig.cap="BMI varying across Countries and daily living activities"}
par(mfrow = c(2, 1))
# Box plots to show how BMI varies across countries
boxplot(bmi ~ country_mod, data = working_data, xlab = "Countries", ylab = "BMI")
# Box plots to show how BMI varies across levels of daily activity
boxplot(bmi ~ adla,
  data = working_data,
  xlab = "Activities of Daily Living index", ylab = "BMI"
)
```


From the boxplots in Figure 1, it is evident that although there isn't a major difference in the value of BMI in each country, there is a difference in the range of values. Austria and Germany seem to have a larger range of values than The Netherlands and Sweden. It is also evident that there is not a large amount of variation in the BMI of those who have high mobility and those who have low mobility. However, it is worth noting that there is a larger range of values of BMI for those who have high mobility.  

\newpage 
The model assumptions of the initial model without separating the data by sex are displayed in Figure 2.

```{r Main-Model, fig.height=5, fig.width=6.5, fig.align='center', fig.cap="Model assumptions for the defined model"}
# Define model formula that applies to both will be applied to both males and females
mod <- bmi ~ . + country_mod:dn004_mod + mother_alive:age + female:chronic_mod +
  female:adla + female:br010_mod + br015_:age + ever_smoked:co007_ +
  ever_smoked:maxgrip + ever_smoked:sphus + hhsize:adla + ch001_:br010_mod +
  adla:age + br010_mod:age

# Define model
fit <- lm(mod, data = working_data)

# Display model assumptions through residuals
par(mfrow = c(2, 2))
plot(fit)
```


The residuals vs fitted plot shows a constant line at zero, justifying the assumption that the expectation is zero. The Normal Q-Q plot is positively skewed, suggesting that the model may be over-fitted. The Scale-Location plot shows only a very slight upwards trend, suggesting that the assumption of a constant variance is not far-fetched. The residuals vs leverage plot shows no data points with high residuals and high leverage, and most of the data points are very close to zero, suggesting that they all have a very similar impact on the model.



```{r model-male-female}
# Define the model formula for each sex
# Note: this model does not include the interaction terms that were in the main
# model as this is applied to each sex separately
sex_mod <- bmi ~ . + country_mod:dn004_mod + mother_alive:age + br015_:age +
  ever_smoked:co007_ + adla:age + br010_mod:age + ever_smoked:maxgrip +
  ever_smoked:sphus + hhsize:adla + ch001_:br010_mod
```

Now, the separate models for males and females may be examined. The summaries for the male and female models are show below.

\newpage

# Model for Males
```{r Sex-model-male, include=FALSE}
# Subset the data to choose only male observations
male_data <- working_data %>%
  filter(female == "No") %>%
  mutate(female = NULL)

# Remove all the NA observations and fit a model to the data
fit_male <- lm(sex_mod, data = na.omit(male_data))

# Perform model selection
better_male_fit <- step(fit_male)

# Round the data in the summary of coefficients to 2 d.p.
sum_better_male <- summary(better_male_fit)
sum_better_male$coefficients <- round(sum_better_male$coefficients, 2)
```


```{r male-summary, out.width="100%"}
# Display summary statistics of model for males
options(width = 300)
sum_better_male
```
 
 
# Model for Females


```{r Sex-model-female, include=FALSE}
# Subset the data to choose only female observations
female_data <- working_data %>%
  filter(female == "Yes") %>%
  mutate(female = NULL)

# Remove all the NA observations and fit a model to the data
fit_female <- lm(sex_mod, data = na.omit(female_data))

# Perform model selection
better_female_fit <- step(fit_female)

# Round the data in the summary of coefficients to 2 d.p.
sum_better_female <- summary(better_female_fit)
sum_better_female$coefficients <- round(sum_better_female$coefficients, 2)
```


```{r female-summary}
# Display summary statistics of model for females
options(width = 300)
sum_better_female
```

From the summaries above, the adjusted $R$-squared values for both models are low, but this is due to variability in the data. After performing model selection procedures, some explanatory variables for both the male and female models were removed, with there now being 18 common variables to both models, 6 and 7 variables being exclusive to males and females respectively.  The relationship between BMI and each of the variables for both males and females is shown in the tables below, where a positive or negative relationship shows that the value of the variable is proportional or inversely proportional respectively to the value of BMI. “Varying for different levels” suggests that each level of the variable may have a different relationship with BMI. 

\begin{table}[!h]
    \centering
    \begin{tabular}{|c|c|c|p{4cm}|p{6cm}|}
    \hline
        \textbf{Variable Name} &  \textbf{Relationship (Male)} & \textbf{Relationship (Female)}\\
        \hline
        mother\_alive  & Negative & Negative \\
        \hline
        ever\_smoked & Negative & Negative \\
        \hline
        iv009\_mod & Positive & Positive \\
        \hline
        hhsize & Negative & Negative \\
        \hline
        chronic\_mod & Positive & Positive \\
        \hline
        adla & Positive & Positive \\
        \hline
        br015\_ & Positive & Positive \\
        \hline
        co007\_ & Negative & Negative \\
        \hline
        br010\_mod & Negative & Negative \\
        \hline
        maxgrip & Positive & Positive \\
        \hline
        sphus & Positive & Positive \\
        \hline
        isced1997\_r & Varying for different levels & Varying for different levels \\
        \hline
        age & Negative & Negative \\
        \hline
        country\_mod & Negative & Varying for different levels \\
        \hline
        ever\_smoked:co007\_ & Positive & Positive \\
        \hline
        br010\_mod:age & Positive & Positive \\
        \hline
        ever\_smoked:maxgrip & Positive & Positive \\
        \hline
        ever\_smoked:sphus & Negative & Negative \\
        \hline
    \end{tabular}
    \label{tab:both}
    \caption{The relationship between BMI and variables common between both sexes.}
\end{table}


\begin{table}[!h]
    \centering
    \begin{tabular}{|c|c|p{4cm}|p{6cm}|}
    \hline
         \textbf{Variable Name} & \textbf{Relationship}\\
         \hline
         partnerinhh & Positive\\
         \hline
         hc029\_ & Positive\\
         \hline
         eurod & Varying for different levels \\
         \hline
         mother\_alive:age & Positive \\
         \hline
         adla:age & Negative\\
         \hline
         hhsize:adla & Positive\\
         \hline 
    \end{tabular}
    \caption{The relationship between BMI and the variables exclusive to males.}
    \label{tab:male}
\end{table}

\begin{table}[!h]
    \centering
    \begin{tabular}{|c|c|p{4cm}|p{6cm}|}
    \hline
         \textbf{Variable Name} & \textbf{Relationship}\\
         \hline
         dn004\_mod & Positive\\
         \hline
         father\_alive & Negative\\
         \hline
         hc012\_ & Negative \\
         \hline
         ch001\_ & Positive \\
         \hline
         dn004\_mod:country\_mod & Negative\\
         \hline
         br015\_:age & Negative\\
         \hline 
         ch001\_:br010\_mod & Negative\\
        \hline
    \end{tabular}
    \caption{The relationship between BMI and the variables exclusive to females.}
    \label{tab:female}
\end{table}

Some variables did not show a very significant relationship with BMI, for example `country_mod`. However, when considering this variable in interactions with other variables, a significant relationship was present. Hence, the country where the interview was conducted may not have made a difference on its own, but it did have a significant impact on other variables such as `dn004_mod`. This was true for other variables as well, including `hhsize`. These variables could be considered as confounding variables.

\newpage
Corresponding to Tables 6,7 and 8, the boxplots below (in Figure 3) help visualise the relationship between BMI and some of the explanatory variables.

```{r boxplot-analysis}
# Plot potential risk factors against BMI for males and females separately

# Number of chronic diseases
chronic_plot <- ggplot(
  na.omit(working_data) %>% select(c(bmi, female, chronic_mod)) %>%
    group_by("female"),
  aes(y = bmi, x = chronic_mod)
) +
  geom_boxplot(aes(fill = female)) +
  xlab("Number of chronic diseases") +
  ylab("BMI") +
  scale_fill_discrete(name = "Sex", labels = c("Male", "Female"))

# Daily living activities
adla_plot <- ggplot(
  na.omit(working_data) %>% select(c(bmi, female, adla)) %>%
    group_by("female"),
  aes(y = bmi, x = adla)
) +
  geom_boxplot(aes(fill = female)) +
  xlab("Activities of daily living") +
  ylab("BMI") +
  scale_fill_discrete(name = "Sex", labels = c("Male", "Female"))

# Drinking Behaviour
br010_plot <- ggplot(
  na.omit(working_data) %>% select(c(bmi, female, br010_mod)) %>%
    group_by("female"),
  aes(y = bmi, x = br010_mod)
) +
  geom_boxplot(aes(fill = female)) +
  xlab("Drinking behaviour") +
  ylab("BMI") +
  scale_fill_discrete(name = "Sex", labels = c("Male", "Female"))

# Maximum grip strength
maxgrip_plot <- ggplot(
  na.omit(working_data) %>% select(c(bmi, female, maxgrip)) %>%
    group_by("female"),
  aes(y = bmi, x = maxgrip)
) +
  geom_boxplot(aes(fill = female)) +
  xlab("Maximum grip strength") +
  ylab("BMI") +
  scale_fill_discrete(name = "Sex", labels = c("Male", "Female"))

# Self-perceived health
sphus_plot <- ggplot(
  na.omit(working_data) %>% select(c(bmi, female, sphus)) %>%
    group_by("female"),
  aes(y = bmi, x = sphus)
) +
  geom_boxplot(aes(fill = female)) +
  xlab("Self-perceived health") +
  ylab("BMI") +
  scale_fill_discrete(name = "Sex", labels = c("Male", "Female"))

# Countries
country_plot <- ggplot(
  na.omit(working_data) %>% select(c(bmi, female, country_mod)) %>%
    group_by("female"),
  aes(y = bmi, x = country_mod)
) +
  geom_boxplot(aes(fill = female)) +
  xlab("Countries") +
  ylab("BMI") +
  scale_fill_discrete(name = "Sex", labels = c("Male", "Female"))
```


```{r present-boxplots-analysis, fig.height=8, fig.width=9, fig.cap="Potential risk factors of obsesity varying with BMI"}

# Present plots together
grid.arrange(chronic_plot, adla_plot, br010_plot, maxgrip_plot, sphus_plot,
  country_plot,
  nrow = 3, newpage = FALSE
)
```


# Conclusion
From the given data, there are some risk factors that affect both males and females, for example smoking habits, level of daily activity, number of chronic diseases, etc. However, some factors exclusively affect males, e.g. depression (EURO-D score) or their age in interaction with their daily activities. On the other hand, some factors exclusively affect females, e.g. number of children, or the amount of vigorous activities in interaction with their age.

\newpage
# Executive Summary
This report illustrates some of the risk factors associated with obesity with BMI as a measure. The data includes participants from four countries - Austria, The Netherlands, Germany and Sweden. When comparing the effect of different risk factors for males and females, it was evident that there is generally a difference in how each factor affects each sex. For example, males with a EURO-D score of 10 or higher were much more likely to have a smaller BMI than those who were less depressed, suggesting that depression may not have much of an impact on their BMI until it becomes significantly severe. On the other hand, there was no apparent relationship between BMI and severity of depression for females at all. The number of chronic illnesses that participants had from each sex seemed to increase their risk of obesity. In addition to this, participants who were generally more active and had a healthier lifestyle, and perceived themselves to be healthier may have a lower risk of being obese. Surprisingly, those who drank regularly had a lower chance of being obese. Participants from Austria and Germany generally had higher values of BMI on average than The Netherlands and Sweden. Being able to make ends meet financially seemed to increase the risk of being obese for both males and females, with the risk being slightly higher for females. Some of this information is visually represented by Figure 4.


```{r exec-plots, fig.height=8, fig.width=9, fig.cap="Possible risk factors of obsesity"}
# Ability to make ends meet
co007_plot <- ggplot(
  na.omit(working_data) %>% select(c(bmi, female, co007_)) %>%
    group_by("female"),
  aes(y = bmi, x = co007_)
) +
  geom_boxplot(aes(fill = female)) +
  xlab("Able to make ends meet") +
  ylab("BMI") +
  scale_fill_discrete(name = "Sex", labels = c("Male", "Female"))

# Present plots for executive summary
grid.arrange(chronic_plot, adla_plot, br010_plot, sphus_plot,
  co007_plot, country_plot,
  nrow = 3
)
```
