------------------------------------------------------------------------

title: "Statistical Case Studies - Project 2 part 1 - Predicting house prices"

author: "Georgia Ettles, Mayez Haris and Niharika Peddinenikalva"

date: "\`r Sys.Date()\`"

output:

pdf_document: default

html_document: default

------------------------------------------------------------------------

This project had equal contributions from all members. All three members sorted and cleaned the data together along with building and selecting the model. Mayez and Georgia refined the models to choose the most suitable one. Niharika and Georgia wrote the code for box plots and other graphs and tables in the project. All three members contributed towards writing the report. 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
set.seed("1234")
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(gridExtra))
suppressPackageStartupMessages(library(countrycode))
suppressPackageStartupMessages(library(visdat))
suppressPackageStartupMessages(library(interactions))
suppressPackageStartupMessages(library(forestplot))
suppressPackageStartupMessages(library(RColorBrewer))
suppressPackageStartupMessages(library(viridis))
suppressPackageStartupMessages(library(xtable))
suppressPackageStartupMessages(library(caret))
suppressPackageStartupMessages(library(naivebayes))
suppressPackageStartupMessages(library(scales))
```

# Load data
```{r data-clean}
# Loading data
data <- read.csv("houseprices.csv")
```


# Logistic regression
## Preprocessing
```{r data-clean}
# Compute the number of missing values
sum(is.na(data[data < 0]))

data$GarageType[which(is.na(data$GarageType) == TRUE)] <- "NA"

# Computing the average SalePrice to make the SalePrice variable binary
lr_data = na.omit(data[, !names(data) == "Id"])

Average_sale_price = mean(data$SalePrice)

# Excluding observations which are negligible in number for certain features
lr_data <- lr_data[!lr_data$Heating %in% c("Floor", "OthW"), ]
lr_data <- lr_data[!lr_data$Utilities %in% c("NoSeWa"), ]
lr_data <- lr_data[!lr_data$Electrical %in% c("Mix"), ]
lr_data <- lr_data[!lr_data$KitchenAbvGr %in% c("3", "0"), ]
lr_data <- lr_data[!lr_data$Neighborhood %in% c("Blueste"), ]
sum(lr_data[,lr_data$SalePrice == Average_sale_price]) # Mention that none of the SP = Average

# Classifying the given prices as below or above average
lr_data[, "SalePrice"] <- ifelse(lr_data[, "SalePrice"] > Average_sale_price, 1, 0)
lr_data[, "SalePrice"] <- factor(lr_data[, "SalePrice"], levels = c(0, 1), 
                                  labels = c("Below Average", "Above Average"))
```



```{r categorising}

# List of variables that are to be categorical and ordinal
categorical_list <- c("Street", "LotShape", "LandContour", "Utilities", "Neighborhood", 
                     "BldgType", "HouseStyle", "RoofStyle", "Foundation", "Heating", 
                     "CentralAir", "Electrical", "KitchenQual", "GarageType", "MoSold")

ordinal_list <- c("FullBath", "HalfBath", "KitchenAbvGr", "Fireplaces")


# Factor the categorical and ordinal variables
for (column in categorical_list) {
  lr_data[, column] <- factor(lr_data[, column])
}

for (column in ordinal_list) {
  lr_data[, column] <- factor(lr_data[, column], ordered = TRUE)
}

summary(lr_data)
```

## Plots and figures
```{r final plots-lr}
### Neighborhood plot ###
# Construct data containing only SalePrice, Neighborhood, and the proportion of houses in each SalePrice category
neighborhood <- lr_data %>% group_by(Neighborhood, SalePrice) %>% summarise(n = n()) %>%
  mutate(n=n/sum(n)) %>%
  ungroup()

# Create a stacked bar graph
neighborhood_his <-ggplot(neighborhood, aes(x = Neighborhood, y = n, fill = SalePrice)) +
  geom_bar(stat = 'identity') + 
  theme(axis.text.y = element_text(size = rel(1.3), hjust = 1), axis.text.x = element_text(size = rel(1.25),hjust = 1), axis.title = element_text(size = rel(1.5))) + 
  labs(x = "Neighborhood", y="Proportion of Houses Sold", fill = "Sale Price") + 
  coord_flip() 

# Save the graph for use in report
suppressMessages(ggsave(file="Neighborhood.pdf", width=6, height=4, dpi=300))


### Overall Quality Plot ###
# Construct data containing only SalePrice, OverallQual, and the propotion of houses in each SalePrice category
Qual <- lr_data %>% 
  group_by(OverallQual,SalePrice)  %>% 
  summarise(n = n()) %>%
  mutate(n=n/sum(n)) %>% 
  ungroup()

# Create a filled line graph
qual_graph <- ggplot(Qual, aes(x = OverallQual, y = n, fill = SalePrice)) +
  geom_bar(stat = 'identity') +
  theme(axis.text.y = element_text(size = rel(1.3), hjust = 1), axis.text.x = element_text(size = rel(1.25)), axis.title = element_text(size = rel(1.5))) + scale_x_continuous(breaks = Qual$OverallQual) +
  labs(x = "Overall Material and Finish Quality",
       y="Proportion of Houses Sold", fill = "Sale Price") + 
  coord_flip() 

# Save the graph for use in report
suppressMessages(ggsave(file="OverallQual.pdf", width=6, height=4, dpi=300))

### House Style Plot ###
# Construct data containing only SalePrice, HouseStyle, and the proportion of houses in each SalePrice category
hstyle <- lr_data %>% group_by(HouseStyle,SalePrice) %>% 
  summarise(n = n()) %>%
  mutate(n=n/sum(n)) %>%
  ungroup()

# Create desired bar graph
hs_graph <- ggplot(hstyle, aes(x =HouseStyle, y = n, fill = SalePrice)) +
  geom_bar(stat = 'identity') + 
  theme(axis.text.y = element_text(size = rel(1.3), hjust = 1), axis.text.x = element_text(size = rel(1.25),hjust = 1), axis.title = element_text(size = rel(1.5))) + 
  labs(x = "Style of Dwelling", y="Proportion of Houses Sold", fill = "Sale Price")  + 
  coord_flip()

# Save the graph for use in report
suppressMessages(ggsave(file="HosueStyle.pdf", width=6, height=4, dpi=300))


### Bedroom Above Grade Plot ###
# Construct data containing only SalePrice, BedroomAbvGr, and the number of houses in each SalePrice category
Bed <- lr_data %>% group_by(BedroomAbvGr,SalePrice)  %>% 
  summarise(n = n()) %>%
  mutate(n=n/sum(n)) %>%
  ungroup()

# construct two graphs showing the different proportions within each sale price
Bed_graph <- ggplot(Bed, aes(x = BedroomAbvGr, y = n, fill = SalePrice)) +
  geom_bar(stat = 'identity', position = position_dodge(), show.legend = FALSE) + facet_grid(. ~SalePrice)  +
  theme(axis.text.y = element_text(size = rel(1.3), hjust = 1), axis.text.x = element_text(size = rel(1.25), hjust = 1), axis.title = element_text(size = rel(1.5))) + 
  labs(y = "Proportion of Houses Sold", x = "No. of Bedrooms Above Basement Level", fill = "Sale Price")

# Save the graph for use in report
suppressMessages(ggsave(file="BedroomAbvGr.pdf", width=6, height=4, dpi=300))
```


```{r variable-plots}

### Neighborhood plot:
# Construct data containing only SalePrice, Neighborhood, and the number of houses in each SalePrice category
neighborhood <- lr_data %>% group_by(Neighborhood, SalePrice) %>% summarise(n = n()) %>%
  mutate(n=n/sum(n)) %>%
  ungroup()


# Create a stacked bar graph
neighborhood_his <-ggplot(neighborhood, aes(x = Neighborhood, y = n, fill = SalePrice)) +
  geom_bar(stat = 'identity') + 
  theme(axis.text.y = element_text(size = rel(1.3), hjust = 1), axis.text.x = element_text(size = rel(1.25),hjust = 1), axis.title = element_text(size = rel(1.5))) + 
  labs(x = "Neighborhood", y="Number of Houses Sold") + 
  coord_flip() 
neighborhood_his
# Save the graph for use in report
#ggsave(file="Neighborhood.pdf", width=8, height=4, dpi=300)
```



## Plots we tested but didn't use
```{r plots not used}
# ### Overall Quality Plot:
# # Construct data containing only SalePrice, OverallQual, and the number of houses in each SalePrice category
# Qual <- lr_data %>% group_by(SalePrice, OverallQual)  %>% 
#   summarise(n = n()) %>%
#   mutate(n=n/sum(n)) %>%
#   ungroup()
# 
# # Create a filled line graph
# qual_graph <- ggplot(Qual, aes(x = OverallQual, y = n)) + 
#   geom_area(aes(fill = SalePrice), alpha = 0.75, position = 'identity', lwd = 0.5, linetype = 1) +
#   geom_line(aes(color = SalePrice), show.legend = FALSE) + 
#   labs(y = "Number of Houses Sold", x = "Overall Material and Finish Quality", fill = "Sale Price") +
#   theme(axis.text.y = element_text(size = rel(1.3), hjust = 1), axis.text.x = element_text(size = rel(1.25)), axis.title = element_text(size = rel(1.5))) 
# 
# 
# qual_graph



# ### Month Sold Plot:
# # Construct data containing only SalePrice, MoSold, and the number of houses in each SalePrice category
# month <- lr_data %>%
#   group_by(MoSold, SalePrice) %>%
#   summarise(n = n()) %>%
#   mutate(n=n/sum(n)) %>%
#   ungroup()
#
# # Create a side-by-side comparison bar graph
# month_his <-ggplot(month, aes(x = MoSold, y = n, fill = SalePrice)) +
#   geom_bar(stat = 'identity') +
#   theme(axis.text.y = element_text(size = rel(1.3), hjust = 1), axis.text.x = element_text(size = rel(1.25)), axis.title = element_text(size = rel(1.5))) +
#   scale_x_discrete(label = month.abb) +
#   labs(x = "Month Sold", y="Proportion of Houses Sold")
# month_his
# # Save the graph for use in report
# #ggsave(file="MoSold.pdf", width=8, height=4, dpi=300)
```


```{r}
# Save the graph for use in report
#ggsave(file="OverallQual.pdf", width=8, height=4, dpi=300)


### House Style Plot:
# Construct data containing only SalePrice, HouseStyle, and the number of houses in each SalePrice category
hstyle <- lr_data %>% group_by(SalePrice, HouseStyle) %>% 
  summarise(n = n()) %>%
  mutate(prop=n/sum(n)) %>%
  ungroup()

# Create a side-by-side bar graph
hs_graph <- ggplot(hstyle, aes(x =reorder(HouseStyle,n), y = prop, fill = SalePrice)) +
  geom_bar(stat = 'identity', position = position_dodge(), show.legend = FALSE) + 
  facet_grid(. ~SalePrice)  +
  geom_text(aes(label = n), fontface = "bold",  vjust = -0.2,  size = 4) + ylim(0,0.6) +
  theme(axis.text.y = element_text(size = rel(1.3), hjust = 1), axis.text.x = element_text(size = rel(1.25), hjust = 1, angle = 45), axis.title = element_text(size = rel(1.5))) + 
  labs(x = "Style of Dwelling", y = "Number of Houses Sold")
hs_graph
# Save the graph for use in report
#ggsave(file="HosueStyle.pdf", width=8, height=4, dpi=300)
```



```{r}
# original formula before model selection
# formula <- SalePrice ~ LotArea + Street + LotShape  + LandContour + Neighborhood + 
#   BldgType + HouseStyle + OverallQual + OverallCond + YearBuilt + RoofStyle + 
#   Foundation + TotalBsmtSF + Heating + CentralAir + Electrical + GrLivArea + 
#   FullBath + HalfBath + BedroomAbvGr + KitchenAbvGr + KitchenQual + TotRmsAbvGrd + 
#   Fireplaces + GarageType + GarageArea + MoSold + YrSold + YearBuilt:OverallCond + 
#   TotRmsAbvGrd:LotArea + HouseStyle:TotRmsAbvGrd + Neighborhood:LotArea +
#   LotArea:BedroomAbvGr


# formula for logistic regression after model selection
formula2 <- SalePrice ~ LotArea + Neighborhood + BldgType + OverallQual + OverallCond + 
  YearBuilt + TotalBsmtSF + CentralAir + GrLivArea + FullBath + BedroomAbvGr + 
  KitchenQual + TotRmsAbvGrd + GarageType + LotArea:BedroomAbvGr + TotRmsAbvGrd:LotArea

```

## Fitting the model and cross validation
```{r}

# Setting up folds for 10-fold cross validation
set.seed(10)
num_folds <- 10 # number of folds

shuffled_data <- lr_data[sample(nrow(lr_data)), ] # shuffle observations

folds <- cut(seq(1, nrow(shuffled_data)), breaks = num_folds, label = FALSE) # create folds

# To store the error rates and brier scores from each fold of cv
error_rates <- numeric(num_folds) 
brier_scores <- numeric(num_folds)
```


```{r}

for (i in 1:num_folds) { # for each fold
  testInd <- which(folds == i, arr.ind = TRUE) # indices of test data
  
  # set test data, training data
  test_data <- shuffled_data[testInd,]
  training_data <- shuffled_data[-testInd,]
  
  # Fit the model using training data and predict using model on test data
  fit <- glm(formula2, data = training_data, family = "binomial")
  predicted <- predict.glm(fit, newdata = test_data, type = "response")
  
  # Compute the Brier scores and error rates
  predicted_brier <- predicted
  predicted <- ifelse(predicted < 0.5, "Below Average", "Above Average")
  predicted <- factor(predicted) 
  true_vals <- test_data$SalePrice
  error_rates[i] <- mean(true_vals != predicted)
  true_vals_brier <- ifelse(true_vals == "Below Average", 0, 1)
  brier_scores[i] <- mean((true_vals_brier - predicted_brier)^2)
}

# compute the average error rates and brier scores from all folds
mean(error_rates)
mean(brier_scores)
```


```{r}
# see model results
summary(fit)
#xtable(summary(fit))

```

## Comparing our model with a model that uses all features
```{r }

# Compare our model with a model using all features except utility
full_formula <- SalePrice ~ LotArea + Street + LotShape + LandContour + Neighborhood + 
  BldgType + HouseStyle + OverallQual + OverallCond + YearBuilt + RoofStyle + 
  Foundation + TotalBsmtSF + Heating + CentralAir + Electrical + GrLivArea + 
  FullBath + HalfBath + BedroomAbvGr + KitchenAbvGr + KitchenQual + TotRmsAbvGrd + 
  Fireplaces + GarageType + GarageArea + MoSold + YrSold
full_error_rates <- numeric(num_folds)
full_brier_scores <- numeric(num_folds)

for (i in 1:num_folds) {
  # identify test indices and perform a train-test split
  testInd <- which(folds == i, arr.ind = TRUE)
  test_data <- shuffled_data[testInd,]
  training_data <- shuffled_data[-testInd,]
  
  # Fit the model using the training data
  full_fit <- glm(full_formula, data = training_data, family = binomial(link='logit'))
  
  # Predict the Sale price probability of being above average on the test set
  predicted <- predict.glm(full_fit, newdata = test_data, type = "response")
  
  # compute the brier scores and error rates
  true_vals <- test_data$SalePrice # for comparing with predicted values
  predicted_brier <- predicted # to compute the brier scores
  predicted <- ifelse(predicted < 0.5, "Below Average", "Above Average")
  predicted <- factor(predicted)  # to compute the error rates
  
  full_error_rates[i] <- mean(true_vals != predicted)
  true_vals_brier <- ifelse(true_vals == "Below Average", 0, 1) 
  full_brier_scores[i] <- mean((true_vals_brier - predicted_brier)^2)
}

mean(full_error_rates)
mean(full_brier_scores)

summary(full_fit) # see full model results
```



# Naive Bayes
```{r}
# create NB data containing SalePrice and 5 chosen features
nb_data <- data[, c("Neighborhood", "HouseStyle", "OverallQual", "BedroomAbvGr", 
                    "MoSold", "SalePrice")]


# convert SalePrice to binary variable in nb_data
Average_sale_price = mean(nb_data$SalePrice)
nb_data[, "SalePrice"] <- ifelse(nb_data[, "SalePrice"] > Average_sale_price, 1, 0)
nb_data[, "SalePrice"] <- factor(nb_data[, "SalePrice"], levels = c(0, 1), 
                                  labels = c("Below Average", "Above Average"))



# factor the qualitative categorical variables
for (column in c("Neighborhood", "HouseStyle", "MoSold")) {nb_data[, column] <- factor(nb_data[, column])}


# defining NB proportions for Neighborhood (multinomial)
counts_neigh <- nb_data %>% group_by(SalePrice) %>% count(Neighborhood)

# add 1 to each count to account for the missing values
counts_neigh$n <- counts_neigh$n + 1


# add new rows that don't exist to account for non-existent data. assign the new counts as 1
new_row1 <- list(SalePrice = 'Above Average', Neighborhood = 'Blueste', n = 1)
new_row2 <- list(SalePrice = 'Above Average', Neighborhood = 'BrDale', n = 1)
new_row3 <- list(SalePrice = 'Above Average', Neighborhood = 'IDOTRR', n = 1)
new_row4 <- list(SalePrice = 'Above Average', Neighborhood = 'MeadowV', n = 1)
new_row5 <- list(SalePrice = 'Below Average', Neighborhood = 'NoRidge', n = 1)
new_row6 <- list(SalePrice = 'Above Average', Neighborhood = 'NPkVill', n = 1)


# add the rows defined above to the dataframe containing counts
counts_neigh <- rbind(counts_neigh, new_row1, new_row2, new_row3, new_row4, new_row5, new_row6)

# add column showing proportions of each category 
probs_df_neigh <- counts_neigh %>% 
  group_by(SalePrice) %>% mutate(prop_neigh = n/sum(n))



# defining NB proportions for HouseStyle (multinomial)
# same process as above
counts_house <- nb_data %>% group_by(SalePrice) %>% count(HouseStyle)

counts_house$n <- counts_house$n + 1

new_row <- list(SalePrice = 'Above Average', HouseStyle = '1.5Unf', n = 1)

counts_house <- rbind(counts_house, new_row)

probs_df_house <- counts_house %>% 
  group_by(SalePrice) %>% mutate(prop_house = n/sum(n))



# defining NB proportions for MoSold (multinomial), same as above
counts_months <- nb_data %>% group_by(SalePrice) %>% count(MoSold)

probs_df_months <- counts_months %>% group_by(SalePrice) %>% mutate(prop_month = n/sum(n))


# defining NB proportions for BedroomAbvGr (multinomial)
counts_bed <- nb_data %>% group_by(SalePrice) %>% count(BedroomAbvGr)

counts_bed$n <- counts_bed$n + 1

bed_row1 <- list(SalePrice = 'Above Average', BedroomAbvGr = 7, n = 1)
bed_row2 <- list(SalePrice = 'Below Average', BedroomAbvGr = 7, n = 1)
bed_row3 <- list(SalePrice = 'Below Average', BedroomAbvGr = 8, n = 1)

counts_bed <- rbind(counts_bed, bed_row1, bed_row2, bed_row3)

probs_df_bed <- counts_bed %>% group_by(SalePrice) %>% mutate(prop_bed = n/sum(n))


# defining proportions for OverallQual (multinomial)

counts_qual <- nb_data %>% group_by(SalePrice) %>% count(OverallQual)

counts_qual$n <- counts_qual$n +1

qual_row1 <- list(SalePrice = 'Above Average', OverallQual = 1, n = 1)
qual_row2 <- list(SalePrice = 'Above Average', OverallQual = 2, n = 1)
qual_row3 <- list(SalePrice = 'Above Average', OverallQual = 3, n = 1)
qual_row4 <- list(SalePrice = 'Below Average', OverallQual = 9, n = 1)

counts_qual <- rbind(counts_qual, qual_row1, qual_row2, qual_row3, qual_row4)

probs_df_qual <- counts_qual %>% group_by(SalePrice) %>% mutate(prop_qual = n/sum(n))

```


```{r badness of poisson for quality and bedrooms}

# this code is used to show that although Poisson seems like a reasonable assumption, it is not actually correct for the provided data


# define the counts dataframe without artifical values
counts_qual2 <- nb_data %>% group_by(SalePrice) %>% count(OverallQual)

# find proportions for each class and category as with the previous dataframes
probs_df_qual2 <- counts_qual2 %>% 
  group_by(SalePrice) %>% mutate(prop_qual = n/sum(n))


# finding means (lambda) for each class for poisson variables to then plot pdf with later on
nb_qual_means <- counts_qual2 %>% group_by(SalePrice) %>% summarise(quality_lamba = mean(OverallQual))

# merge the lambdas with the count and proportion df
qual_df <- merge(probs_df_qual2, nb_qual_means, by='SalePrice')

# calculate value from poisson pdf for each count
qual_df <- qual_df %>% group_by(SalePrice) %>% mutate(poiss = dpois(OverallQual, lambda = quality_lamba))


# plot the poisson pdf on top of original data to verify the distribution
qual_graph_pois <- ggplot(qual_df, aes(x = OverallQual, y = prop_qual, fill = SalePrice)) + facet_grid(. ~SalePrice) + geom_bar(stat = 'identity', show.legend = FALSE) + geom_line(aes(y=poiss)) +
  geom_text(aes(label = n), vjust = -0.23,  size = 3.2) +
  theme(axis.text.y = element_text(size = rel(1.3), hjust = 1), axis.text.x = element_text(size = rel(1.25)), axis.title = element_text(size = rel(1.5))) +
  labs(x = "Overall Material and Finish Quality", y="Proportion of Houses Sold")
qual_graph_pois



# same process as above, but for no. of bedrooms
counts_bed2 <- nb_data %>% group_by(SalePrice) %>% count(BedroomAbvGr)

probs_df_bed2 <- counts_bed2 %>% group_by(SalePrice) %>% mutate(prop_bed = n/sum(n))

nb_bed_means <- counts_bed2 %>% group_by(SalePrice) %>% summarise(bed_lamba = mean(BedroomAbvGr))

bed_df <- merge(probs_df_bed2, nb_bed_means, by='SalePrice')

bed_df <- bed_df %>% group_by(SalePrice) %>% mutate(poiss = dpois(BedroomAbvGr, lambda = bed_lamba))

bed_graph_pois <- ggplot(bed_df, aes(x = BedroomAbvGr, y = prop_bed, fill = SalePrice)) +
  geom_bar(stat = 'identity', position = position_dodge(), show.legend = FALSE) + 
  geom_line(aes(y=poiss), stat = 'identity') +
  facet_grid(. ~SalePrice)  +
  geom_text(aes(label = n),  vjust = -0.2,  size = 4) + 
  theme(axis.text.y = element_text(size = rel(1.3), hjust = 1), axis.text.x = element_text(size = rel(1.25), hjust = 1), axis.title = element_text(size = rel(1.5))) + labs(y = "Proportion of Houses Sold", x = "Number of Bedrooms Above Basement Level")


# save plots
bed_graph_pois
suppressMessages(ggsave(file="bed_graph_pois.pdf", width=9, height=6, dpi=300))

qual_graph_pois
suppressMessages(ggsave(file="qual_graph_pois.pdf", width=9, height=6, dpi=300))

```


```{r NB proportion plots}
# # proportion plot for Neighborhood with counts
# neigh_plot_nb <- ggplot(probs_df_neigh, aes(x = Neighborhood, y = prop_neigh, fill = SalePrice)) +
#   geom_bar(stat = 'identity') + facet_grid(. ~SalePrice) + ylim(0,0.26) +
#   geom_text(aes(label = n), hjust = -0.2,  size = 3.2) +
#   theme(axis.text.y = element_text(size = rel(1.3), hjust = 1), axis.text.x = element_text(size = rel(1.25),hjust = 1), axis.title = element_text(size = rel(1.5))) + 
#   labs(x = "Neighborhood", y="% of Houses Sold") + 
#   coord_flip() 



# proportion plot for HouseStyle
house_plot_nb <- ggplot(probs_df_house, aes(x =reorder(HouseStyle,n), y = prop_house, fill = SalePrice)) +
  geom_bar(stat = 'identity', position = position_dodge(), show.legend = FALSE) + 
  facet_grid(. ~SalePrice)  +
  geom_text(aes(label = n),  vjust = -0.2,  size = 4) + ylim(0,0.6) +
  theme(axis.text.y = element_text(size = rel(1.3), hjust = 1), axis.text.x = element_text(size = rel(1.25), hjust = 1, angle = 45), axis.title = element_text(size = rel(1.5))) + 
  labs(x = "Style of Dwelling", y = "Proportion of Houses Sold")


# # proportion plot for MoSold
# month_plot_nb <- ggplot(probs_df_months, aes(x = MoSold, y = prop_month, fill = SalePrice)) + facet_grid(. ~SalePrice) +
#   geom_bar(stat = 'identity') +
#   geom_text(aes(label = n), hjust = -0.2,  size = 3.2)  + ylim(0, 0.205) +
#   theme(axis.text.y = element_text(size = rel(1.3), hjust = 1), axis.text.x = element_text(size = rel(1.25)), axis.title = element_text(size = rel(1.5))) +
#   scale_x_discrete(labels=month.abb) +
#   labs(x = "Month Sold", y="Proportion of Houses Sold") + coord_flip()


# proportion plot for OverallQual

qual_plot_nb <- ggplot(probs_df_qual, aes(x = OverallQual, y = prop_qual, fill = SalePrice)) + facet_grid(. ~SalePrice) +
  geom_bar(stat = 'identity') +
  geom_text(aes(label = n), vjust = -0.23,  size = 3.2) +
  theme(axis.text.y = element_text(size = rel(1.3), hjust = 1), axis.text.x = element_text(size = rel(1.25)), axis.title = element_text(size = rel(1.5))) +
  labs(x = "Overall Material and Finish Quality", y="Proportion of Houses Sold")


# proportion plot for BedroomAbvGr
bed_plot_nb <- ggplot(probs_df_bed, aes(x = BedroomAbvGr, y = prop_bed, fill = SalePrice)) + facet_grid(. ~SalePrice) +
  geom_bar(stat = 'identity') +
  geom_text(aes(label = n), vjust = -0.23,  size = 3.2) +
  theme(axis.text.y = element_text(size = rel(1.3), hjust = 1), axis.text.x = element_text(size = rel(1.25)), axis.title = element_text(size = rel(1.5))) +
  labs(x = "No. of Bedrooms Above Basement Level", y="Proportion of Houses Sold")



# save all plots
# neigh_plot_nb
# suppressMessages(ggsave(file="neigh_plot_nb.pdf", width=9, height=6, dpi=500))

house_plot_nb
suppressMessages(ggsave(file="house_plot_nb.pdf", width=6, height=4, dpi=300))


# month_plot_nb
# suppressMessages(ggsave(file="month_plot_nb.pdf", width=6, height=4, dpi=300))


qual_plot_nb
suppressMessages(ggsave(file="qual_plot_nb.pdf", width=6, height=4, dpi=300))


bed_plot_nb
suppressMessages(ggsave(file="bed_plot_nb.pdf", width=6, height=4, dpi=300))


house_plot_nb
qual_plot_nb
bed_plot_nb
```


```{r NB model assessment}
# Cross Validation and fitting the model
set.seed(10)
num_folds <- 10 # number of folds for cv

shuffled_data <- nb_data[sample(nrow(nb_data)), ] # shuffle data

folds <- cut(seq(1, nrow(shuffled_data)), 
             breaks = num_folds, label = FALSE) # create folds

# vectors to store error rates, brier scores and log scores
error_rates_nb <- numeric(num_folds) 
brier_scores_nb <- numeric(num_folds)
log_scores_nb <- numeric(num_folds)


# Perform 10-fold cross validation
for (i in 1:num_folds) { # for each fold
  testInd <- which(folds == i, arr.ind = TRUE) # test indices for fold i
  
  # assign test data, training data with features and response separately
  test_data <- data.matrix(shuffled_data[testInd, -6]) 
  training_data <- data.matrix(shuffled_data[-testInd, -6])
  training_response <- shuffled_data$SalePrice[-testInd]
  test_labels <- shuffled_data$SalePrice[testInd] # store test labels
  true_vals <- ifelse(test_labels == "Below Average", 0, 1) # test responses as numeric
  
  # fit the naive bayes model which uses 5 multinomial features from the training data
  nb_fit <- multinomial_naive_bayes(x = training_data, y=training_response)
  # use model to predict on test data
  predicted <- predict(nb_fit, newdata = test_data, type = "prob") 
  
  
  # compute error rates
  predicted_labels <- ifelse(predicted[,1] > 0.5, "Below Average", "Above Average")
  error_rates_nb[i] <- mean(test_labels != predicted_labels)
  
  # compute brier scores
  predicted_brier <- predicted[, 2]
  brier_scores_nb[i] <- mean((true_vals - predicted_brier)^2)
  
  # compute log scores
  predicted_logscore <- ifelse(true_vals == 0, predicted[,1], predicted[,2])
  log_scores_nb[i] <- mean((true_vals - predicted_logscore)^2)
  
}

# compute all three mean average scores
mean(error_rates_nb)
mean(brier_scores_nb)
mean(log_scores_nb)

# # Check the probability coefficients for the features
# xtable(nb_fit$params)

# Present all the cross validation scores in a table for the report
error_rates_all <- c(mean(error_rates), mean(full_error_rates), mean(error_rates_nb))
brier_scores_all <- c(mean(brier_scores), mean(full_brier_scores), mean(brier_scores_nb))
log_scores_all <- c(NA, NA, mean(log_scores_nb))
scores_df <- data.frame("Error Rates" = error_rates_all, "Brier scores" = brier_scores_all,
           row.names = c("Logistic regression model (Final)", 
                         "Logistic regression model (Full)", "Naive Bayes model"))


nb_fit # see model results

```