```{r Packages & Setup}
knitr::opts_chunk$set(echo = FALSE)
suppressPackageStartupMessages(library("tidyverse"))
suppressPackageStartupMessages(library("dplyr"))
suppressPackageStartupMessages(library("gridExtra"))
suppressPackageStartupMessages(library("visdat"))
suppressPackageStartupMessages(library("countrycode"))
suppressPackageStartupMessages(library("ggplot2"))
suppressPackageStartupMessages(library("knitr"))
suppressPackageStartupMessages(library("nullabor"))
suppressPackageStartupMessages(library("ggthemes"))
suppressPackageStartupMessages(library("broom"))
suppressPackageStartupMessages(library("latex2exp"))
suppressPackageStartupMessages(library("reshape2"))
suppressPackageStartupMessages(library("readxl"))
suppressPackageStartupMessages(library("lubridate"))
suppressPackageStartupMessages(library("openintro"))
suppressPackageStartupMessages(library("janitor"))
```

```{r INITIAL CLEANING}

solutions <- c("8", "9", "6", "2", "7", "2", "6", "5")

orig_data <- read_excel("lineup_survey_results.xlsx", sheet = "cleaned_data")

orig_data <- orig_data %>% 
  clean_names() %>% 
  filter(lu1 == solutions[1] & lu5 == solutions[5])


# factor columns
to_factor <- c("stats_y2", "level_of_study", "year_of_study",  
               "lu1", "lu1_confidence", "lu2", "lu2_confidence", "lu3", "lu3_confidence", "lu4", "lu4_confidence", 
               "lu5", "lu5_confidence", "lu6", "lu6_confidence", "lu7", "lu7_confidence", "lu8", "lu8_confidence")

for (col in to_factor) {
  orig_data[[col]] <- factor(orig_data[[col]])
}


# remove comments columns
data <- subset(orig_data, select = -c(lu1_comments, lu2_comments, lu3_comments, lu4_comments, 
                                      lu5_comments, lu6_comments, lu7_comments, lu8_comments))

data <- data %>% 
  mutate(lu1_correct = if_else(lu1 == solutions[1], TRUE, FALSE),
         lu2_correct = if_else(lu2 == solutions[2], TRUE, FALSE),
         lu3_correct = if_else(lu3 == solutions[3], TRUE, FALSE),
         lu4_correct = if_else(lu4 == solutions[4], TRUE, FALSE),
         lu5_correct = if_else(lu5 == solutions[5], TRUE, FALSE),
         lu6_correct = if_else(lu6 == solutions[6], TRUE, FALSE),
         lu7_correct = if_else(lu7 == solutions[7], TRUE, FALSE),
         lu8_correct = if_else(lu8 == solutions[8], TRUE, FALSE))

correctness_data <- subset(data, select = c(lu1_correct, lu2_correct, lu3_correct, lu4_correct, 
                                            lu5_correct, lu6_correct, lu7_correct, lu8_correct))
```

Need to create bar graphs or pie charts to show how many participants responded to a question correctly. Do mention the first and fifth lineup as well, but without any graphs. Additionally, state how not too many entries had to be deleted. √

Need some visuals on who has done Stats Y2, on the level of study (UG or PG), and on the year of study (only need this for UG as PG only has one year anyway).

Create a function to evaluate the p-value from each lineup using the formula from the literature. √


```{r p-values of lineups}

lineup_p_val <- function(lineup_number, size = 9) {
  lineup <- correctness_data[[lineup_number]]
  observers <- length(lineup)
  num_correct <- length(lineup[lineup == TRUE])
  p_val <- 0
  for (i in num_correct:observers) {
    p_val <- p_val + choose(observers, i) * ((1/size)^i) * ((size-1)/size)^(observers-i)
  }
  round(p_val, 6)
}

p_values <- c()
for (lineup_number in 1:8) {
  p_values <- append(p_values, lineup_p_val(lineup_number))
}

p_values_df <- data.frame("lineup" = 1:8,
                          "p_value" = p_values)

p_vals_nullabor <- c()
for (col in names(correctness_data)) {
  p_vals_nullabor <- append(p_vals_nullabor, 
                            nullabor::pvisual(x = sum(correctness_data[[col]] == TRUE), K = 55, m = 9)[[3]])
}

p_values_df <- cbind(p_values_df, p_vals_nullabor)
p_values_df$diff <- abs(p_values_df$p_value - p_values_df$p_vals_nullabor)

```

```{r correctness per question (setting up dataframes)}

colnames(correctness_data) <- c("Lineup 1", "Lineup 2", "Lineup 3", "Lineup 4", 
                                "Lineup 5", "Lineup 6", "Lineup 7", "Lineup 8")

reshaped_correctness_df <- tidyr::gather(correctness_data, key = "Lineup", value = "Correct.Answer")

correct_count_df <- reshaped_correctness_df %>%
  group_by(Lineup, Correct.Answer) %>%
  summarise(count = n(), .groups = "keep") %>%
  spread(Correct.Answer, count, fill = 0)


# convert data to long format for easier plotting
correct_prop_df <- tidyr::gather(correct_count_df, key = "Correct.Answer", value = "Count", -Lineup)

# Calculate proportion for each lineup
correct_prop_df <- correct_prop_df %>%
  group_by(Lineup) %>%
  mutate(Proportion = Count / sum(Count))

```


```{r correctness stacked bar graph}

# Plot horizontal bar graph
correct_prop_bar_graph <- ggplot(correct_prop_df, aes(x = Proportion, y = Lineup, fill = Correct.Answer)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_text(aes(label = ifelse(Proportion > 0, scales::percent(Proportion), "")),
            position = position_stack(vjust = 0.5), size = 4, color = "white") +
  labs(title = "Proportion of Correct and Incorrect Counts for Each Lineup",
       x = "Proportion",
       y = "Lineup") +
  scale_fill_manual(values = c("TRUE" = "seagreen", "FALSE" = "firebrick3")) +
  scale_x_continuous(labels = scales::percent_format(scale = 100)) +
  scale_y_discrete(limits=rev) +
  theme_minimal()

```

```{r correctness pie charts}

correct_pie_charts <- ggplot(correct_prop_df, aes(x = "", y = Proportion, fill = Correct.Answer)) +
  geom_bar(stat = "identity", position = "stack", width = 1) +
  geom_text(aes(label = ifelse(Proportion > 0, scales::percent(Proportion), "")),
            position = position_stack(vjust = 0.5), size = 4, color = "white") +
  coord_polar(theta = "y") +
  facet_wrap(~ Lineup, nrow = 2, ncol = 4) +
  labs(title = "Proportion of Correct and Incorrect Counts for Each Lineup",
       x = NULL,
       y = NULL) +
  scale_fill_manual(values = c("TRUE" = "seagreen", "FALSE" = "firebrick3")) +
  theme_minimal() +
  theme(panel.grid = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank())

```


